A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes

### 2017-10-30 paper/aosgflp.tex
Incorporate edits to paper

The reliability of many engineered products follow a similar pattern. Relatively high rates of failure occurring early (``infant mortality'') due to manufacturing defects.  After this ``burn-in'' period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to spurious inferences about a product's reliability and suboptimal decisions. This highlights the importance of choosing a sufficiently flexible model.

@@

Nonparametric methods require relatively few assumptions, but they are not suitable for prediction. Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be useful. Appropriate parametric models can make better use of available data. For inference, likelihood-based methods are well-equipped to deal with truncated and/or censored data. Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare populations of interest, borrow strength across groups, and incorporate prior information.

Often no information is available to identify why a unit failed, even when the failed units are available. Physical failure mode analysis (sometimes referred to as ``autopsy'') can be extremely time consuming and expensive. This presents a dilemma for analysts: the data may indicate multiple failure modes, contraindicating the use of a unimodal parametric distribution. However, without any knowledge of cause of failure, traditional competing risk models are not identifiable. \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model. This model can provide a solution to the problem of unknown cause of failure (masking) in the special case where there is evidence of two modes of failure which impact different stages of product life. It avoids non-identifiability by introducing a parameter representing the population proportion defective. Where this parameter is zero, the GLFP model reduces to a unimodal distribution. \\

The GLFP model has a meaningful parameterization and accommodates lifetime data with infant mortality and wearout.  Unfortunately, it can require a lot of data to fit due to the model's complexity. When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes ``soft'' constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a ``pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.

The structure of the paper is as follows. Section~\ref{sec:Background} introduces the proposed method and potential applications and examines previous related work in the literature.  Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of hard drive failure data from the company, Backblaze.  In Section~\ref{sec:Data}, we describe the complete Backblaze Hard Drive data set, which is used throughout as a motivating example. In Section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations. Section~\ref{sec:Data analysis} applies the hierarchical GLFP model to the hard disk drive data. First, we do model selection among four models of increasing complexity, using approximate leave-one-out cross-validation. Next, for the selected model, we assess the model fit using a graphical comparison of replicated data sets generated from the posterior distribution to the observed data. Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime. In Section~\ref{sec:Discussion}, we suggest potential applications of our model and consider limitation for inference implied by the model assumptions.

@@

The GLFP model can be understood as a marginalized mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$. $\delta_i$ is an indicator for whether or not an observational unit is defective. Expressed conditional on $\delta_i$,
\begin{align*}
P(T\le t | \delta_i=1) =& 1 -(1-F_1(t))(1-F_2(t))\\
P(T\le t | \delta_i=0) =& F_2(t).
\end{align*}

@@

Taking the derivative of the (marginal) cdf, the density for the GLFP is
\begin{equation}
h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + f_2(t|t_{p2},\sigma_2)\left(1-\pi F_2(t|t_{p2},\sigma_2\right).
\end{equation}

@@

\subsubsection{Likelihood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time; that is, the first time reported for the unit.  Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$). The likelihood for the GLFP is a function of $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is given by
\begin{equation*}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation*}

\section{Motivating Example}
\label{sec:Data}
\subsection{Backblaze hard drive data}
Backblaze is a company that offers cloud backup storage to protect against data loss.  Since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods.  Drives are run until failure.  When a hard drive fails it is permanently removed and replaced.  In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.  Every quarter Backblaze makes its data publicly available through their website \citep{backblaze}. In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.  No other analysis or modeling of the failure data is provided.  Backblaze does, however, encourage others to further analyze its data.

@@

\noindent These prior distributions put probability mass on a wide range of values for all model parameters, much larger than we would expect for a typical Weibull distribution.  Thus, we consider the priors to be relatively uninformative.

Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for Drive-model 14.  The parameter $\pi$ is an estimate of the proportion of drives susceptible to to infant mortality.  As expected, this fraction is small with a median value of 0.05.  The parameter $\pi$ estimates for the two Weibull distributions also match intuition.  The infant mortality failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.  Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, implying an increasing hazard function.  The two quantiles are also well identified with the infant mortality failure mode having an much earlier time to reach the 0.50 quantile compared to the time to reach the 0.20 quantile for the wearout mode.

@@

Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.  After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with pointwise confidence bands provides a visual check for distributional goodness of fit.  We applied this method using the Kaplan-Meier nonparametric estimate of the empirical cdf.  With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \citet[Chapter 11]{meeker} (see Appendix for a detailed description).

In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted K-M estimate with axes on the Weibull probability scale.  The plot also contains 90\% pointwise uncertainty bands associated with each estimate. While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to capture the acceleration of the hazard rate between 8000 and 20,000 hours. %The discrepancy around 30,000 suggests that the GLFP may overestimate the proportion failing in the far right tail of the distribution.  While this discrepancy may appear troubling, we suggest that the right tail is of less importance for electronic goods due to obsolescence.

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{km14-prob-plus}
  \caption{Estimated GLFP for Drive-model 14 plotted on Weibull paper.  Dashed line corresponds to the median of the posterior draws; pointwise 90\% credible band is in red.  Solid line is adjusted K-M estimate with pointwise 90\% credible band in blue.}
  \label{ex1-overlay}
\end{figure}

@@

In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population-specific parameters hierarchically, borrowing strength across the sub-populations.  Let
\begin{equation}
y_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$.  Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{h(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-H(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $t_{ig}$ is the end of the observed lifetime of unit $i$ in group $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.  All times are again in hours.

Next, we model these parameters themselves as random variables, varying across the different sub-populations.  Let
\begin{equation*}
\sigma_{1g} \ind \op{Log-normal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
\end{equation*}
\begin{equation*}
\sigma_{2g} \ind \op{Log-normal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{T}\left(0, 1\right)
\end{equation*}
\begin{equation}
\label{eq:hier-model}
t_{p_{1}g} \equiv \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
\end{equation}
\begin{equation*}
t_{p_{2}g} \equiv \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
\end{equation*}
\begin{equation*}
\pi_g \ind \op{Logit-normal}(\eta_\pi, \tau^2_\pi).
\end{equation*}

As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of log quantile and scale since the data we encounter are more informative about the left side of their respective lifetime distribution. We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $T(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (e.g., $\beta_{2g} = 1/\sigma_{2g}>1$). We choose to model these parameters with log-normal and logit-normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.

@@

\section{Analysis of the Backblaze hard drive data}
\label{sec:Data analysis}
\subsection{Modeling}
\subsubsection{Hierarchical models}
We considered four models, all based on (\ref{eq:hier-model}), which are identified by which model parameters are held constant across drive-models. These are (from most to least restrictive):

@@

The set of model specifications was chosen based data, interpretation of the model, as well as estimation considerations.  Product specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.  Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.  Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by drive-model.

For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive-models.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).

@@

For Models 1-4, different sets of restrictions required different prior distribution specifications, which were assigned as follows:

\paragraph{Model 1.} Constrain all drive-models to the same GLFP distribution. For this ``reduced" model we assume the same prior distributions as in Example 1.

\paragraph{Model 2.} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
$$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
$$\sigma_1 \sim \op{Log-normal}\langle 1.4 \times 10^{-1},\, 7.1\rangle,$$ 
$$t_{p1} \sim \op{Log-normal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$

\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.

\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.

@@

\subsection{Computation}
\label{sec:Computation}
Each model was fit using the {\tt
  rstan} \citep{rstan} package in {\tt R} \citep{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
\citep{betancourt}. HMC jointly updates all model parameters by simulating energy preserving paths with random initial momentums along the posterior density. This is done to reduce autocorrelation and efficiently explore the posterior. Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1, 2 and 3 for a total of 6,000 post burn-in iterations and 16 chains were run for Model 4 for a total of 24,000 post burn-in iterations. The Gelman-Rubin potential scale reduction factor was used to provide a check for adequate mixing of the multiple chains. Upon convergence, $\hat{R}$, converges to 1.  The  respective maximum values of $\hat{R}$ across all model parameters for Models 1 through 4 (5, 50, 95 and 140 parameters, respectively) were 1.005, 1.028, 1.009 and 1.004. Other diagnostics provided by the software (tree-depth, and divergent transitions) did not indicate problems with sampling. Plots of the posterior draws were inspected for parameters with the fewest effective samples; these did not suggest features that were inadequately explored.

@@

Figure \ref{fig:fig2} provides a visual comparison for Models 1 through 4 by plotting the pointwise posterior median of time to failure. The large left panel shows the adjusted K-M estimate of the same. The adjusted K-M estimates on the right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit. The smaller plots correspond to Models 1 through 4.  Note that while all the paths for the posterior estimates necessarily pass through the origin, the K-M estimates do not, because, for the K-M estimator, inference is entirely conditional on survival up to left-most left-trunctation time. To make the plot comparable, we use an adjustment based on a parametric model (Model 4).

%Should we be brief discussing the plot?
These plots suggest that Model 1 is insufficient to explain the observed data. Model 2 certainly captures more of the heterogeneity observed in the K-M estimates; we observe that the assumption of a common $\sigma_2(=\sigma_{2,g})$ results in a very similar progressions in the rate of failure, which may be too restrictive. Model 3, which removes this restriction on $\sigma_{2,g}$ suggests that there is evidence to support variation in this parameter among drive-models. While Model 3 and 4 do display differences, they are relatively subtle.

\begin{figure}[H]
\includegraphics[width=\textwidth]{heterogeneity-compare}
\caption{Left: K-M estimates of the time to failure for each of the drive-models in the Backblaze data. Right: Pointwise posterior median time to failure curves for Models 1-4, ordered left to right, top to bottom.}
\label{fig:fig2}
\end{figure}

We can also compare posterior estimates of time to failure under Models 1-4 for each drive-model individually.  For these comparisons, we calculate, pointwise, the posterior median of the proportion failing over a grid. That is, for every model, and for a fixed set, $\tilde{t}=\{\tilde{t}_1,\ldots,\tilde{t}_M\}$, we compute 
\begin{equation}
\label{pointwise-medians}
\op{median}\left\{\op{H}\left(\tilde{t}_m|\pi_{g}^{(s)},t_{p_{1}g}^{(s)},\sigma_{1g}^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)}\right);s=1,\ldots,S \right\},
\end{equation}
where $\theta^{(s)}$ is the $s^{th}$ posterior sample from $p(\theta|y)$.
%Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.

@@

There are two sources of variation at play: the
posterior uncertainty in the parameters, which largely depends on the
sample size, and the uncertainty in future
observations conditional on the parameters. The {\em posterior
  predictive} distribution incorporates both sources.
\begin{equation*}
  p(t_{g,new}|t) = \int p(t_{g,new}|\theta_g) p(\theta_g|t_g) \, d\theta_g
\end{equation*}

>>>>

The reliability of many engineered products follows a similar pattern. Relatively high rates of failure occurring early (``infant mortality'') is due to manufacturing defects.  After this ``burn-in'' period, failure rates stabilize after the majority of defective units have failed and are no longer in the population.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to spurious inferences about a product's reliability and suboptimal decisions. This highlights the importance of choosing a sufficiently flexible model.

@@

Nonparametric methods require relatively few assumptions, but may not be suitable for prediction when extrapolation is required. Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be useful. Appropriate parametric models can make better use of available data. For inference, likelihood-based methods are able to deal with truncated and/or censored data. Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare populations of interest, borrow strength across groups, and incorporate prior information.

Often no information is available to identify why a unit failed, even when the failed units are available. Physical failure mode analysis (sometimes referred to as ``autopsy'') can be extremely time consuming and expensive. This presents a dilemma for analysts: the data may indicate multiple failure modes, contraindicating the use of a unimodal parametric distribution. However, without any knowledge of cause of failure, traditional competing risk models are not identifiable. \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model. This model can provide a solution to the problem of unknown cause of failure (masking) in the special case where there is evidence of two modes of failure which impact different stages of product life. It avoids non-identifiability by introducing a parameter representing the population proportion defective. When this parameter is zero, the GLFP model reduces to a unimodal distribution. \\

The GLFP model has a meaningful parameterization and accommodates lifetime data with both infant mortality and wearout.  Unfortunately, it can require a lot of data to fit due to the model's complexity. When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes ``soft'' constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a ``pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.

The structure of the paper is as follows. Section~\ref{sec:Background} introduces the proposed method and potential applications and examines previous related work.  Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of hard drive failure data from the company, Backblaze.  In Section~\ref{sec:Data}, we describe the complete Backblaze Hard Drive data set, which is used throughout as a motivating example. In Section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations. Section~\ref{sec:Data analysis} applies the hierarchical GLFP model to the hard disk drive data. First, we do model selection among four models of increasing complexity, using approximate leave-one-out cross-validation. Next, for the selected model, we assess the model fit using a graphical comparison of replicated data sets generated from the posterior distribution to the observed data. Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime. Section~\ref{sec:Discussion} describes potential applications of our model and consider limitation for inference implied by the model assumptions.

@@

The GLFP model can be understood as a mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$. $\delta_i$ is an indicator for whether or not a unit $i$ is defective. Expressed conditional on~$\delta_i$,
\begin{align*}
P(T\le t | \delta_i=1) =& 1 -(1-F_1(t))(1-F_2(t))\\
P(T\le t | \delta_i=0) =& F_2(t).
\end{align*}

@@

Taking the derivative of the (marginal) cdf, the density for the GLFP is
\begin{align*}
h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_2(t|t_{p2},\sigma_2\right).
\end{align*}

@@

\subsubsection{Likelihood}
We now give the general form for the likelihood function, taking into account left truncation and right censoring.  Let $t_{i}$ denote the end of the observed lifetime of unit $i$, in hours.  Let $t_i^L$ be the left truncation time, the age of unit $i$ when data reporting commenced. Additionally, let $c_i$ be an indicator for censoring; $c_i=1$ if the failure time is right-censored, $c_1=0$ if the unit failed (at time $t_i$). The likelihood for the GLFP is a function of the parameter $\bm{\theta} = (\pi,t_{p1}, \sigma_1, t_{p2}, \sigma_2)$.  Assuming the lifetimes of all units are independent, the likelihood for the data, $t_1,\ldots,t_n$ is given by
\begin{equation*}
L(\bm{\theta})= \prod_{i=1}^{n} \left[\frac{h(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})}\right]^{1-c_i} \left[ \frac{1-H(t_i;\bm{\theta})}{1-H(t_i^L;\bm{\theta})} \right]^{c_i}
\end{equation*}

\section{Motivating Example}
\label{sec:Data}
\subsection{Backblaze hard drive data}
Backblaze is a company that offers cloud backup storage to protect against data loss.  Since 2013 it has been collecting data on hard drives operating at its facility.  The purpose is to provide consumers and businesses with reliability information on different hard drive-models.  The hard drives continuously spin in controlled storage pods.  Drives are run until failure.  When a hard drive fails it is removed and replaced.  In addition, the number of storage pods is increasing as Backblaze adds drives to its storage capacity.  Every quarter Backblaze makes its hard drive data publicly available through their website (\url{https://www.backblaze.com/b2/hard-drive-test-data.html}, Accessed April 1, 2016)\nocite{backblaze}. In addition, Backblaze publishes summary statistics of the different hard drive-models currently operating.  No other analysis or modeling of the failure data is provided.  Backblaze does, however, encourage others to further analyze their data.

@@

\noindent These prior distributions put probability mass on a wide range of values for all model parameters --- much larger than we would expect for a typical Weibull distribution.  Thus, we consider these prior distributions to be relatively uninformative.

Table \ref{table:1} gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for Drive-model 14.  The parameter $\pi$ is an estimate of the proportion of drives susceptible to infant mortality.  As expected, this proportion is small with a median value of 0.05.  The shape parameter estimates for the two Weibull distributions also match intuition.  The infant mortality failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.  Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, implying an increasing hazard function.  The two quantiles are also well identified with the infant mortality failure mode having an much earlier time to reach the 0.50 quantile compared to the time to reach the 0.20 quantile for the wearout mode.

@@

Probability plotting is a simple method to assess and compare the adequacy of members of the log-location-scale families of distributions.  After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with pointwise confidence bands provides a visual check for distributional goodness of fit.  We applied this method using the Kaplan-Meier nonparametric estimate of the empirical cdf.  With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \citet[Chapter 11]{meeker} (see Appendix for a detailed description).

In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted K-M estimate with axes on the Weibull probability scale.  The plot also contains 90\% pointwise uncertainty bands associated with each estimate. While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to adequately describe the rapid increase in the empirical cdf between 8000 and 20,000 hours. %The discrepancy around 30,000 suggests that the GLFP may overestimate the proportion failing in the far right tail of the distribution.  While this discrepancy may appear troubling, we suggest that the right tail is of less importance for electronic goods due to obsolescence.

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{km14-prob-plus}
  \caption{The estimated GLFP for Drive-model 14 plotted on Weibull paper.  The dashed line corresponds to the median of the posterior draws; pointwise 90\% credible bounds are shown in red.  The solid line is an adjusted K-M estimate with pointwise 90\% credible band in blue.}
  \label{ex1-overlay}
\end{figure}

@@

In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population-specific parameters hierarchically, borrowing strength across the sub-populations.  Let
\begin{equation}
y_{ig} \ind \op{GLFP}\left( \pi_g, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP is a function of the parameters $\bm{\theta_g} = (\pi_{g}, t_{p_{1}g}, \sigma_{1g}, t_{p_{2}g}, \sigma_{2g})$.  Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{h(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-H(t_{ig};\bm{\theta_g})}{1-H(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $t_{ig}$ is the observed failure or survival time of unit $i$ in group $g$; $t_{ig}^L$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.  All times are again in hours.

In the hierarchical model, the parameters are modeled as random variables, varying across the different sub-populations.  Let
\begin{equation*}
\sigma_{1g} \ind \op{Log-normal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
\end{equation*}
\begin{equation*}
\sigma_{2g} \ind \op{Log-normal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{T}\left(0, 1\right)
\end{equation*}
\begin{equation}
\label{eq:hier-model}
t_{p_{1}g} \equiv \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
\end{equation}
\begin{equation*}
t_{p_{2}g} \equiv \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Log-normal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
\end{equation*}
\begin{equation*}
\pi_g \ind \op{Logit-normal}(\eta_\pi, \tau^2_\pi).
\end{equation*}

As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of a quantile and the shape parameter because the data we encounter are more informative about the lower tail of their respective lifetime distribution (note that if $T$ has a Weibull distribution, $\sigma$ is a shape parameter for the distribution of $T$ and a scale parameter for the distribution of $\log(T)$). We truncate the distribution of $\sigma_{2g}$ at $1$ (indicated by $T(0,1)$ above) restricting the wearout failure mode to have an increasing hazard function (e.g., $\beta_{2g} = 1/\sigma_{2g}>1$). %We choose to model these parameters with log-normal and logit-normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.

@@

\section{Analysis of the Backblaze hard drive data}
\label{sec:Data analysis}
\subsection{Modeling}
\subsubsection{Hierarchical models}
We considered four models, all based on (\ref{eq:hier-model}), differing by which model parameters are held constant across drive-models. These are (from most to least restrictive):

@@

The set of model specifications was chosen based data, interpretation of the model, as well as estimation considerations.  Drive-model-specific parameters for the the wearout failure mode $(t_{p_{2}g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across drive-models in the right tails of the failure distribution.  Going from a common model for all product brands and gradually increasing the complexity of the model, we conclude that Model 4 provides the best description of the data.  Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the wearout failure mode to vary by drive-model.

For all of the models we consider, the parameters for the infant mortality failure mode are held in common across drive-models.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$). Also the failure-time distribution of the defective subpopulation is not of high interest (as opposed to the proportion defective.)

@@

For Models 1--4, different sets of restrictions required different prior distribution specifications, which were assigned as follows:

\paragraph{Model 1} Constrain all drive-models to the same GLFP distribution. For this ``reduced" model we assume the same prior distributions as in Example 1.

\paragraph{Model 2} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
$$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
$$\sigma_1 \sim \op{Log-normal}\langle 1.4 \times 10^{-1},\, 7.1\rangle,$$ 
$$t_{p1} \sim \op{Log-normal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$

\paragraph{Model 3} $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Prior distributions for the constrained parameters are the same as for Model 2.

\paragraph{Model 4} $\pi_g$, $t_{p,g}$ and $\sigma_{2g}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.

@@

\subsection{Computation}
\label{sec:Computation}
Each model was fit using the {\tt
  rstan} \citep{rstan} package in {\tt R} \citep{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
\citep{betancourt}. HMC jointly updates all model parameters by simulating energy preserving paths with random initial momentums along the posterior density. This is done to reduce autocorrelation and efficiently explore the posterior. Multiple chains were run, each with 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1, 2 and 3 for a total of 6,000 post burn-in iterations and 16 chains were run for Model 4 for a total of 24,000 post burn-in iterations. The Gelman-Rubin potential scale reduction factor was used to provide a check for adequate mixing of the multiple chains. Upon convergence, $\hat{R}$, converges to 1.  The  respective maximum values of $\hat{R}$ across all model parameters for Models 1 through 4 (5, 50, 95 and 140 parameters, respectively) were 1.005, 1.028, 1.009 and 1.004. Other diagnostics provided by the software (tree-depth, and divergent transitions) did not indicate problems with sampling. Plots of the posterior draws were inspected for parameters with the fewest effective samples; these did not suggest features that were inadequately explored.

@@

Figure \ref{fig:fig2} provides a visual comparison for Models 1 through 4 by plotting the pointwise posterior median of the cdf for each drive-model.  To make the plot comparable, we use an adjustment based on a parametric model (Model 4). The large left panel shows the adjusted K-M estimate of the same. The adjusted K-M estimates on the right side of Figure \ref{fig:fig2} shows that the set of estimated failure curves for all four models that we fit. The smaller plots correspond to Models 1 through 4.  Note that while all the paths for the posterior estimates necessarily pass through the origin, the K-M estimates do not, because, for the K-M estimator, inference is entirely conditional on survival up to left-most left-trunctation time.

%Should we be brief discussing the plot?
These plots suggest that Model 1 is insufficient to explain the observed data. Model 2 certainly captures more of the heterogeneity observed in the K-M estimates; we observe that the assumption of a common $\sigma_2(=\sigma_{2,g})$ results in a very similar progressions in cdf, which may be too restrictive. Model 3, which removes this restriction on $\sigma_{2,g}$ suggests that there is evidence to support variation in this parameter among drive-models. While Model 3 and 4 do display differences, they are relatively subtle.

\begin{figure}[H]
\includegraphics[width=\textwidth]{heterogeneity-compare}
\caption{Left: Adjusted K-M estimates of the time to failure for each of the drive-models in the Backblaze data. Right: Pointwise posterior median time to failure curves for Models 1-4, ordered left to right, top to bottom.}
\label{fig:fig2}
\end{figure}

We can also compare posterior estimates of time to failure under Models 1-4 for each drive-model individually.  For these comparisons we calculate pointwise the posterior median of the proportion failing over a grid. That is, for every model, and for a fixed set, $\tilde{t}=\{\tilde{t}_1,\ldots,\tilde{t}_M\}$, we compute 
\begin{equation}
\label{pointwise-medians}
\op{median}\left\{\op{H}\left(\tilde{t}_m|\pi_{g}^{(s)},t_{p_{1}g}^{(s)},\sigma_{1g}^{(s)},t_{p_{2}g}^{(s)},\sigma_{2g}^{(s)}\right);s=1,\ldots,S \right\},
\end{equation}
where $\theta^{(s)}$ is the $s^{th}$ posterior sample from $p(\theta|y)$.
%Unlike maximum likelihood, by including the hyperparameters in the probability model, there is a data-dependent penalization that results in shrinkage toward a common overall model.

@@

There are two sources of variation at play: the
posterior uncertainty in the parameters, which largely depends on the
amount of data, and the uncertainty in future
observations conditional on the parameters. The {\em posterior
  predictive} distribution incorporates both sources of variability.
\begin{equation*}
  p(t_{g,new}|t) = \int p(t_{g,new}|\theta_g) p(\theta_g|t_g) \, d\theta_g
\end{equation*}
